Joint maximum-likelihood of phylogeny and ancestral states is not consistent

### 2019-03-29 versions/response.v2.txt

The reviewer is correct that this story is slightly different than the one in the previous version of the paper. Regarding the bootstrap, we remind the reviewer that we are interested in the case of "infinite data," and thus the bootstrap doesn't apply. We have also taken out the speculation about higher likelihoods on the wrong topology.

> Getting complete results about topological consistency is feasible.
> The arguments that led to the closed form solution for the limiting
> edge-length estimates for tau_1 can be repeated similarly for tau2.
> This would allow, for any given x* & y*, comparison of the limiting
> lnl per site for tau1 vs tau2. Whichever topology has the larger lnl
> is the long-run estimated topology. In cases where hat w=1 for both
> topologies, conclusions are different but definite. Thus it seems a
> relatively straightforward exercise given the work the authors have
> already done (I hope I am not missing a subtle difficulty) to have a
> computer determine for a grid of (x*,y*) whether (a) hat w=1 exactly
> for both topologies, (b) the difference in limiting lnl over
> topologies is either positive or negative or (c) hat w neq 1 for at
> least one of the topologies and the difference in limiting lnl is 0 up
> to machine precision. I would emphasize that the cases (a) and (b)
> constitute rigorous proof. The cases (c) likely are rare and
> correspond to boundary cases, so the boundary may have some slight
> uncertainty. There may be some aesthetic appeal to getting analytical
> results rather than using a computer but being able to draw clear
> conclusions about strict topological inconsistency is important.

In our revised work we found no evidence of topological inconsistency, and as mentioned above we have removed such speculation from the paper. "Machine precision" is not involved: the optimal value of w is exactly equal to 1 in the limit.

@@

> The language used to refer to the approach and references to the
> literature are awkward. The references to profile likelihood work
> suggest that there is a theoretical basis for the approach (that paper
> suggests consistency) but the settings considered in such work are
> quite different, I believe. Integrated likelihood usually refers to
> something different too. Finally, as Reviewer 3 makes clear the only
> current implementation of the approach was intended as an
> approximation to ML rather than as ML. The central issue is whether
> treating unobserved data (ancestral states here) as parameters
> sometimes works; the EM algorithm were developed with the idea that
> this is not what should be done. In any case, references directly tied
> to that issue would be more relevant. Additional suggestions include:
> 1) Make a clear statement in the intro about why the usual likelihood
> is considered the right likelihood: it is the probability of the
> *observed* data. 2) The point that large numbers of parameters need
> not defeat likelihood methods is valid but it should be made clear
> when referencing such work that the setting is different.

We use "integrated likelihood" in the same manner as Goldman. We have replaced "profile likelihood" with "joint objective function" throughout, unless we were specifically making a comparison with profile methods. We have removed references that are unrelated to the approaches considered in the manuscript, and defer these to the discussion for the interested reader to follow up with if they so choose. Thanks.

Thank you for your suggestion in point 1; we have added it to the introduction.
We have also clarified in regards to point 2 that the joint objective function is related to profile likelihoods, but is not the same thing.
